NGO analysis

### 2022-07-31 Dashboard/report.Rmd
Deleting commented-out code.

## Function to read Excel file -------------------------------------------------
widen_excel <- function(file) {
  read_excel(file) |>
    # uniting Category and Name to deal with multiple 'Other'
    unite(
      entry, c(Category, Name),
      sep = "_"
    ) |>
    # deleting Sub-category and Firm
    select(-`Sub-category`, -Firm) |>
    # pivoting to create 'tidy' variables
    pivot_wider(names_from = entry, values_from = Value) |>
    # cleaning names
    clean_names()
}
### Reading html ---------------------
htmldata <- read_proff(
  params$htmlfile,
  as.character(params$start), as.character(params$end)
)

@@

## Currency scale-------------------------------------------------------------
# Function to check for currency and use currency code or ?
currency_scale <-
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(htmldata$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(htmldata$valutakode[1], " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )

## Currency caption-----------------------------------------------------------
# Real currency caption
currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency.\n"
    )

# Fake currency caption - turns the values around
# currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
#   "Note: not all values are denominated in the same currency.", ""
# )
```

@@

```{r revenue_expenses_plot, fig.asp=1, fig.show="hold", out.width="45%", fig.cap="Based on the categories \`Revenue' and \`Expenses' in the Excel sheet."}
## Revenue--------------------------------------------------------------------
revenue_plot <-
  # testing for data quality
  if (datatest == TRUE) {
    # printing 'bad data' plot if data are bad
    NA
  } else {
    exceldata |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        `Private contributions` = revenue_private_contributions,
        `Membership fees` = revenue_membership_fees,
        `Public grants` = revenue_public_grants,
        Investments = revenue_investments,
        `Operational activities` = revenue_operational_activities,
        `Commercial activities` = revenue_commercial_activities,
        Other = revenue_other
      ) |>
      # Pivoting to use fill
      pivot_longer(
        c(`Private contributions`:Other),
        names_to = "Revenue source"
      ) |>
      # Ordering for the plot
      mutate(
        `Revenue source` =
          ordered(`Revenue source`,
            levels =
              c(
                "Other", "Commercial activities",
                "Operational activities", "Investments",
                "Membership fees", "Private contributions",
                "Public grants"
              )
          )
      ) |>
      ggplot(aes(Year, value, fill = `Revenue source`)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum)
      ) +
      # Specifying the number of values, as the base palette only includes 6
      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
      labs(x = "", y = "", title = "Revenue (in thousands)") +
      theme_light() +
      theme(title = element_text(size = 14))
  }

>>>>

# Creating a data frame for currency (for some reason the functions in the
# dashboard create bad data, but *only* in a *parameterised* R Markdown report).
read_proff_currency <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill(
      valutakode,
      .direction = "downup"
    )
}

## Function to read Excel file -------------------------------------------------
widen_excel <- function(file) {
  read_excel(file) |>
    # uniting Category and Name to deal with multiple 'Other'
    unite(
      entry, c(Category, Name),
      sep = "_"
    ) |>
    # deleting Sub-category and Firm
    select(-`Sub-category`, -Firm) |>
    # pivoting to create 'tidy' variables
    pivot_wider(names_from = entry, values_from = Value) |>
    # cleaning names
    clean_names()
}
### Reading html ---------------------
# Creating the main data frame
htmldata <- read_proff(
  params$htmlfile,
  as.character(params$start), as.character(params$end)
)

# Creating the currency data frame
html_currency <- read_proff_currency(params$htmlfile,
  as.character(params$start), as.character(params$end))

@@

## Currency scale-------------------------------------------------------------
# Function to check for currency and use currency code or ?
currency_scale <-
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(html_currency$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(unique(html_currency$valutakode), " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )

## Currency caption-----------------------------------------------------------
# Real currency caption
currency_caption <- ifelse(length(unique(html_currency$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency.\n"
    )
```

@@

```{r revenue_expenses_plot, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("Based on the category \`Revenue' in the Excel sheet.", "Based on the category \`Expenses' in the Excel sheet.")}
## Revenue--------------------------------------------------------------------
revenue_plot <-
  # testing for data quality
  if (datatest == TRUE) {
    # printing 'bad data' plot if data are bad
    NA
  } else {
    exceldata |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        `Private contributions` = revenue_private_contributions,
        `Membership fees` = revenue_membership_fees,
        `Public grants` = revenue_public_grants,
        Investments = revenue_investments,
        `Operational activities` = revenue_operational_activities,
        `Commercial activities` = revenue_commercial_activities,
        Other = revenue_other
      ) |>
      # Pivoting to use fill
      pivot_longer(
        c(`Private contributions`:Other),
        names_to = "Revenue source"
      ) |>
      # Ordering for the plot
      mutate(
        `Revenue source` =
          ordered(`Revenue source`,
            levels =
              c(
                "Other", "Commercial activities",
                "Operational activities", "Investments",
                "Membership fees", "Private contributions",
                "Public grants"
              )
          )
      ) |>
      ggplot(aes(Year, value, fill = `Revenue source`)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum)
      ) +
      # Specifying the number of values, as the base palette only includes 6
      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
      labs(x = "", y = "", title = "Revenue (in thousands)") +
      theme_light() +
      theme(title = element_text(size = 14))
  }
